---
title: "Week 5 Report"
author: "Leo Tesler"
format: html
---

## GitHub Link

[https://github.com/leotesler/autoresearch_mlb_project_stat_390](https://github.com/leotesler/autoresearch_mlb_project_stat_390)

## Experiment Log

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(DT)

results <- read_csv(here("autoresearch/results.csv"))

results |> 
  datatable()
```

## Metric Plot

```{r}
results |> 
  mutate(j_index = as.numeric(j_index)) |> 
  ggplot(aes(x = iteration, y = j_index)) +
  geom_line() +
  labs(
    title = "Change in J Index Over Time",
    x = "Iteration #",
    y = "J Index"
  ) +
  geom_vline(xintercept = 38, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = 59, linetype = "dashed", color = "blue", size = 1) +
  annotate("text", x = 39, y = 0.835, label = "End of Loop 1", 
           hjust = -0.1, color = "red") +
  annotate("text", x = 48, y = 0.81, label = "End of Loop 2", 
           hjust = -0.1, color = "blue") +
  annotate("point", x = 1, y = 0.817, size = 3, color = "yellow") +
  annotate("text", x = 1, y = 0.82, label = "Baseline") +
  annotate("point", x = 58, y = 0.839, size = 3, color = "green") +
  annotate("text", x = 58, y = 0.841, label = "Best") +
  theme_classic()
```
## Keep/Discard/Crash Summary

```{r}
results |> 
  mutate(
    best = cummax(lag(j_index, default = 0)),
    improvement = j_index - best,
    category = case_when(
      improvement >= 0.001 ~ "keep",
      j_index > 0 ~ "discard",
      TRUE ~ "crash"
    )
  ) |> 
  select(iteration, j_index, category) |> 
  datatable()
```

## What Actually Worked

  - SMOTE with tuned lasso over the iteration-31 feature set: 0.827846
  - Upper-minors defensive spectrum/range features: improved to 0.828994
  - Log ISO/HR/FB transforms + Low-A → High-A progression deltas: large jump to 0.836954
  - High-A → AA progression deltas: improved to 0.838531
  - Rookie → Low-A progression deltas: tiny improvement to final best, 0.838587
  
## What Didn't Work

  - Raw fielding variables were noisy; removing them helped stabilize the recipe.
  - Career-wide defensive spectrum/progression features dropped to 0.832840.
  - Position-specific defensive share interactions dropped to 0.833152.
  - Downsampling the best feature set dropped badly to 0.820807.
  - SMOTE with neighbors = 3 dropped to 0.834896.
  
## Final Model

  The final model:

  - Keeps the iteration-31 upper-minors offensive/age-adjusted features.
  - Adds log transforms for AA/AAA ISO, HR, and FB%.
  - Adds Rookie→Low-A, Low-A→High-A, High-A→AA, and AA→AAA progression deltas.
  - Adds upper-minors defensive spectrum/range/share features.
  - Removes raw inn_*, putouts_*, and assists_* fielding columns after deriving defensive features.
  - Uses step_smote(success) and logistic_reg(penalty = tune(), mixture = 1).